% In this file you should put the actual content of the blueprint.
% It will be used both by the web and the print version.
% It should *not* include the \begin{document}
%
% If you want to split the blueprint content into several files then
% the current file can be a simple sequence of \input. Otherwise It
% can start with a \section or \chapter for instance.

\part{Brownian motion}

\paragraph{Overview}

This part of the blueprint is a guide to the formalization of a standard Brownian motion in Lean using Mathlib. There are two main parts to this formalization:
\begin{itemize}
  \item a development of the theory of Gaussian distributions, the construction of a projective family of Gaussian distributions and its projective limit by the Kolmogorov extension theorem,
  \item a proof of a Kolmogorov-Chentsov continuity theorem, following \cite{kratschmer2023kolmogorov}.
\end{itemize}

Putting the two sides together, we then build a stochastic process that fits the definition of a Brownian motion on the real line.

\paragraph{Status} The formalization is complete.

\paragraph{Formalization authors} Rémy Degenne, Markus Himmel, David Ledvinka, Etienne Marion, Peter Pfaffelhuber.

With additional contributions from Jonas Bayer, Lorenzo Loccioli, Pietro Monticone, Alessio Rondelli and Jérémy Scanvic.

\input{chapters/characteristic_function.tex}
\input{chapters/process.tex}
\input{chapters/gaussian.tex}
\input{chapters/projective_family.tex}
\input{chapters/kolmogorov_chentsov.tex}
\input{chapters/brownian.tex}


\part{Stochastic integral}

\paragraph{Overview}

We describe the construction of a stochastic integral.

\paragraph{Status} The formalization is ongoing.

\paragraph{Formalization authors} Anyone is welcome to contribute!

\input{chapters/debut.tex}
\input{chapters/doob_meyer.tex}
\input{chapters/stochastic_integral.tex}

\bibliographystyle{amsalpha}
\bibliography{bib}
