\chapter{Modifications with cadlag paths}

\section{Cadlag modifications of martingales}

\subsection{Upcrossings}


\begin{lemma}\label{lem:tendsto_of_no_upcrossings}
  \mathlibok
  \lean{tendsto_of_no_upcrossings}
Let $u : \beta \to \alpha$, for $\alpha$ a densely ordered, conditionally complete linear order equipped with the order topology.
Let $S$ be a dense subset of $\alpha$ and $f$ a filter of $\beta$.
If for all $a < b$ in $S$, the number of upcrossings of the interval $[a, b]$ by $u$ along $f$ is finite, then $u$ tends to a limit along $f$.
\end{lemma}

\begin{proof}\leanok

\end{proof}


\begin{lemma}\label{lem:upcrossing_simpleProcess_le}
  \uses{def:elemStochIntegral, def:simpleProcess}
Let $X : T \to \Omega \to \mathbb{R}$ be a stochastic process on a finite time domain $T$.
Then for all $a < b$ in $\mathbb{R}$ and $t \in T$, there exists an elementary predictable set $A$ such that the number of upcrossings $U_t[a, b]$ of the interval $[a, b]$ before time $t$ satisfies
\begin{align*}
  (b - a) U_t[a, b] \le (\mathbb{1}_A \bullet X)_t + \max\{a - X_t, 0\}
  \:.
\end{align*}
\end{lemma}

\begin{proof}

  See \href{https://almostsuremath.com/2009/12/06/upcrossings-downcrossings-and-martingale-convergence/}{this chapter} of \cite{almostsuremath}.

  In the proof, note that $A$ can be written as a finite disjoint union of sets of the form
  $\{(t, \omega) \mid \sigma(\omega) < t \le \tau(\omega)\}$ for stopping times $\sigma, \tau$.
  To see that each is an elementary predictable set, note that if $T = \{t_0, \ldots, t_n\}$ where
  $t_0 < \cdots < t_n$, then it can be written as a finite disjoint union of sets
  \begin{align*}
    \{\sigma(\omega) < t \le \tau(\omega)\} &= \bigcup_{k=1}^{n} {t_k} \times \{\sigma < t_k \le \tau\}
    \\ &= \bigcup_{k=1}^{n} (t_{k-1}, t_k] \times \{\sigma \le t_{k-1} < \tau\}
  \end{align*}
  where $\{\sigma \le t_{k-1} < \tau\} \in \mathcal{F}_{t_{k-1}}$.
  Therefore $A$ is an elementary predictable set.

\end{proof}



\subsection{Cadlag modifications}


\begin{theorem}\label{thm:exists_rightContinuous_modification_of_bounded_elemStochIntegral}
  \uses{def:elemStochIntegral, def:modification}
  \leanok
  \lean{ProbabilityTheory.exists_modification_left_right_limit}
Let $X : T \to \Omega \to \mathbb{R}$ be an adapted stochastic process such that $X$ is integrable and for every $t \in T$ the set $\{\mathbb{E}[(\mathbb{1}_A \bullet X)_t] \mid A \text{ elementary predictable}\}$ is bounded.
% \begin{enumerate}
%   \item $X$ is integrable and for every $t \in T$ the set $\{\mathbb{E}[(\mathbb{1}_A \bullet X)_t] \mid A \text{ elementary predictable}\}$ is bounded.
%   \item For every $t \in T$ the set $\{(\mathbb{1}_A \bullet X)_t \mid A \text{ elementary predictable}\}$ is bounded in probability.
% \end{enumerate}
Then $X$ has a modification $Y$ which has left and right limits everywhere and such that there is a countable set $S \subseteq T$ for which $Y$ is right-continuous on $T \setminus S$.
\end{theorem}

\begin{proof}
  \uses{lem:upcrossing_simpleProcess_le, lem:tendsto_of_no_upcrossings}
  Let $D$ be a countable dense subset of $T$ (e.g.\ the rationals in $T$) and let $t \in T$ and let
  $S$ be a finite subset of $D \cap [0, t]$. By Lemma~\ref{lem:upcrossing_simpleProcess_le}, for all
  $a < b$ in $\mathbb{R}$, the number of upcrossings $U_S[a, b]$ of the interval $[a, b]$ along $S$
  satisfies
  \begin{align*}
    (b - a) U_S[a, b] \le (\mathbb{1}_A \bullet X)_t + \max\{a - X_t, 0\}
  \end{align*}
  for some elementary predictable set $A$.
  Taking expectations and using the boundedness hypothesis, we obtain
  \begin{align*}
    (b - a) \mathbb{E}[U_S[a, b]] \le L + \mathbb{E}[\max\{a - X_t, 0\}] \le L + |a| + \mathbb{E}[|X_t|]
  \end{align*}
  where $L$ is a bound for $\{\mathbb{E}[(\mathbb{1}_A \bullet X)_t] \mid A \text{ elementary predictable}\}$.

  Now, let $S_1 \subseteq S_2 \subseteq \cdots$ be an increasing sequence of finite subsets of
  $D \cap [0, t]$ with $\bigcup_n S_n = D \cap [0, t]$.
  The number of upcrossings $U_{S_n}[a, b]$ is increasing in $n$.
  By monotone convergence,
  \begin{align*}
    (b - a) \mathbb{E}[U_{D \cap [0, t]}[a, b]] = \lim_{n \to \infty} (b - a) \mathbb{E}[U_{S_n}[a, b]]
      \le L + |a| + \mathbb{E}[|X_t|] < \infty.
  \end{align*}
  In particular, $U_{D \cap [0, t]}[a, b] < \infty$ a.s. for each pair $a < b$.
  Since $\mathbb{Q}^2$ is countable, we conclude that almost surely, for all rational $a < b$, the
  number of upcrossings of $[a, b]$ by $X$ along $D \cap [0, t]$ is finite.

  By Lemma~\ref{lem:tendsto_of_no_upcrossings}, finite upcrossings of all rational intervals implies
  that the left and right limits of $s \mapsto X_s(\omega)$ along $D$ exist at every $t \in T$, for
  almost every $\omega$. Define
  $$\tilde{Y}_t(\omega) = \lim_{\substack{s \downarrow t \\ s \in D}} X_s(\omega)$$
  whenever this limit exists, and $\tilde{Y}_t(\omega) = 0$ otherwise.
  Then $\tilde{Y}$ has left and right limits everywhere (along $T$) almost surely, and is
  right-continuous at every point of $D$.

  We claim that the set
  \begin{align*}
    S = \{t \in T \mid \tilde{Y}_t \ne X_t \text{ with positive probability}\}
  \end{align*}
  is countable.
  For each $n \in \mathbb{N}$, consider the set
  \begin{align*}
    S_n = \{t \in T \cap [0, n] \mid \mathbb{P}(|\tilde{Y}_t - X_t| > 1/n) > 1/n\}
    \:.
  \end{align*}
  We have $S = \bigcup_n S_n$, so it suffices to show each $S_n$ is finite.
  If $S_n$ were infinite, it would contain a monotone sequence $t_k$ converging to some limit
  $t^* \in [0, n]$. Adding $\{t_k\}$ to $D$ and repeating the upcrossing argument, we would obtain
  that $X_{t_k}$ converges along $D \cup \{t_k\}$ almost surely. But $\tilde{Y}_{t_k}$ also converges
  (since $\tilde{Y}$ has left/right limits), giving $\tilde{Y}_{t_k} - X_{t_k} \to 0$ in probability,
  contradicting the definition of $S_n$. Therefore each $S_n$ is finite and $S$ is countable.

  Define
  \begin{align*}
    Y_t = \begin{cases} X_t & \text{if } t \in S \:, \\ \tilde{Y}_t & \text{if } t \notin S \:. \end{cases}
  \end{align*}
  Then $Y_t = X_t$ almost surely for all $t$: for $t \in S$ this is by definition, and for
  $t \notin S$ this follows from $\tilde{Y}_t = X_t$ a.s. Thus $Y$ is a modification of $X$.

  Since $\tilde{Y}$ has left and right limits everywhere almost surely, and $Y$ differs from
  $\tilde{Y}$ only on the countable set $S$, $Y$ also has left and right limits everywhere almost surely.
  Furthermore, $Y$ agrees with $\tilde{Y}$ on $T \setminus S$, so $Y$ is right-continuous on $T \setminus S$.
\end{proof}

In the following, we say that a stochastic process $X$ is \emph{right-continuous in probability} if for all $t \in T$, $\lim_{s \downarrow t} X_s = X_t$ in which the limit is taken in probability.

\begin{corollary}\label{cor:exists_caldag_modification}
  \uses{def:elemStochIntegral, def:modification}
  \leanok
  \lean{ProbabilityTheory.exists_modification_isCadlag}
Let $X : T \to \Omega \to \mathbb{R}$ be an adapted stochastic process which is right-continuous in probability and such that the boundedness condition of Theorem~\ref{thm:exists_rightContinuous_modification_of_bounded_elemStochIntegral} holds.
Then $X$ has a cadlag modification.
\end{corollary}

\begin{proof}\leanok
  \uses{thm:exists_rightContinuous_modification_of_bounded_elemStochIntegral}
Let $Y$ be the modification of $X$ given by Theorem~\ref{thm:exists_rightContinuous_modification_of_bounded_elemStochIntegral}.
$Y$ has left and right limits everywhere and there is a countable set $S$ such that $Y$ is right-continuous on $T \setminus S$. It remains to show that $Y$ is right-continuous at the points of $S$.
Let $Y_{t+}$ denote the right limit of $Y$ at $t$.
Since $X$ is right-continuous in probability, for all $t \in S$, $Y_t = Y_{t+}$ almost surely.
Therefore, since $S$ is countable, we have that almost surely $Y_t = Y_{t+}$ for all $t \in S$ (and hence for all $t \in T$).
$Y$ is therefore almost surely cadlag.
We can modify $Y$ on the null set to obtain a cadlag modification of $X$.
\end{proof}


\begin{lemma}\label{lem:Submartingale.integral_elemStochIntegral_bounded}
  \uses{def:elemStochIntegral, def:Submartingale}
Let $X : T \to \Omega \to \mathbb{R}$ be a submartingale.
Then for every $t \in T$, the set $\{\mathbb{E}[(\mathbb{1}_A \bullet X)_t] \mid A \text{ elementary predictable}\}$ is bounded.
\end{lemma}

\begin{proof}
  \uses{lem:submartingale_iff_integral_elemStochIntegral_nonneg, cor:Submartingale.integral_elemStochIntegral_le}
Since $X$ is a submartingale, for any elementary predictable set $A$, we have (Corollary~\ref{cor:Submartingale.integral_elemStochIntegral_le})
\begin{align*}
  0 \le \mathbb{E}[(\mathbb{1}_A \bullet X)_t] \le \mathbb{E}[X_t - X_0]
  \:.
\end{align*}
As $X_t$ and $X_0$ are integrable, the set $\{\mathbb{E}[(\mathbb{1}_A \bullet X)_t] \mid A \text{ elementary predictable}\}$ is bounded.
\end{proof}


\begin{lemma}\label{lem:Submartingale.exists_cadlag_modification_of_rightContinuous}
  \uses{def:modification, def:Submartingale}
Let $X : T \to \Omega \to \mathbb{R}$ be a submartingale which is right-continuous in probability.
Then $X$ has a cadlag modification.
\end{lemma}

\begin{proof}
  \uses{cor:exists_caldag_modification, lem:Submartingale.integral_elemStochIntegral_bounded}
We apply Corollary~\ref{cor:exists_caldag_modification}, in which the boundedness condition of Theorem~\ref{thm:exists_rightContinuous_modification_of_bounded_elemStochIntegral} holds by Lemma~\ref{lem:Submartingale.integral_elemStochIntegral_bounded}.
\end{proof}


\begin{lemma}\label{lem:Submartingale.exists_modifications_limits}
  \uses{def:modification, def:Submartingale}
Let $X : T \to \Omega \to \mathbb{R}$ be a submartingale.
Then $X$ has a modification $Y$ such that for all $t \in T$, $Y$ has left and right limits at $t$ and such that there is a countable set $S \subseteq T$ for which $Y$ is right-continuous on $T \setminus S$.
\end{lemma}

\begin{proof}
  \uses{thm:exists_rightContinuous_modification_of_bounded_elemStochIntegral, lem:Submartingale.integral_elemStochIntegral_bounded}
Apply Theorem~\ref{thm:exists_rightContinuous_modification_of_bounded_elemStochIntegral}, in which the boundedness condition holds by Lemma~\ref{lem:Submartingale.integral_elemStochIntegral_bounded}.
\end{proof}


\begin{lemma}\label{lem:Submartingale.uniformIntegrable_of_antitone_of_ge}
  \uses{def:Submartingale}
Let $X : T \to \Omega \to \mathbb{R}$ be a submartingale and let $t_n$ be a decreasing sequence in $T$ which is bounded below.
Then the family $\{X_{t_n}\}_{n \in \mathbb{N}}$ is uniformly integrable.
\end{lemma}

\begin{proof}

\end{proof}


\begin{lemma}\label{lem:Submartingale.le_condExp_of_tendsto}
  \uses{def:Submartingale}
Let $X : T \to \Omega \to \mathbb{R}$ be a submartingale and $t \in T$.
Let $t_n$ be a decreasing sequence in $T$ converging to $t$, and such that $X_{t_n}$ tends to a limit $X_{t+}$ almost surely.
Then $X_t \le P[X_{t+} \mid \mathcal{F}_t]$ almost surely.
\end{lemma}

\begin{proof}
  \uses{lem:Submartingale.uniformIntegrable_of_antitone_of_ge}
By the submartingale property, for all $n$ we have that $X_t \le P[X_{t_n} \mid \mathcal{F}_t]$ almost surely.
By Lemma~\ref{lem:Submartingale.uniformIntegrable_of_antitone_of_ge}, the family $\{X_{t_n}\}_{n \in \mathbb{N}}$ is uniformly integrable.


TODO: conclude that $P[X_{t+} \mid \mathcal{F}_t]$ is the limit of $P[X_{t_n} \mid \mathcal{F}_t]$.

\end{proof}


\begin{theorem}\label{thm:Submartingale.exists_cadlag_modification_iff_rightContinuous}
  \uses{def:modification, def:Submartingale, def:rightContinuous}
Let $X : T \to \Omega \to \mathbb{R}$ be a submartingale with respect to a right-continuous filtration.
Then $X$ has a cadlag modification if and only if $t \mapsto \mathbb{E}[X_t]$ is right-continuous.
\end{theorem}

\begin{proof}
  \uses{lem:Submartingale.exists_modifications_limits, lem:Submartingale.le_condExp_of_tendsto, lem:Submartingale.uniformIntegrable_of_antitone_of_ge}
Let $Y$ be the modification of $X$ given by Lemma~\ref{lem:Submartingale.exists_modifications_limits}.
$Y$ has left and right limits everywhere and there is a countable set $S$ such that $Y$ is right-continuous on $T \setminus S$.
It remains to show that for each $t \in S$, $Y$ is right-continuous at $t$ almost surely. We will then be able to modify $Y$ on a null set to obtain a cadlag modification of $X$.
Let $t \in S$ and let $Y_{t+}$ denote the right limit of $Y$ at $t$.

We show that $Y_t = P[Y_{t+} \mid \mathcal{F}_t]$ almost surely, and that $P[Y_{t+} \mid \mathcal{F}_t] = Y_{t+}$ almost surely, which will conclude the proof.
For the second equality it suffices to show that $Y_{t+}$ is $\mathcal{F}_t$-measurable, which follows from the right-continuity of the filtration.

For the first equality, it suffices to show that $P[Y_{t+} \mid \mathcal{F}_t] - Y_t$ is an almost surely nonnegative random variable with zero expectation.
Nonnegative follows from Lemma~\ref{lem:Submartingale.le_condExp_of_tendsto}.
The expectation is $P[Y_{t+}] - P[Y_t]$. Let $t_n$ be a decreasing sequence in $T$ converging to $t$.
By right-continuity of $t \mapsto \mathbb{E}[Y_t]$, we have that
\begin{align*}
  P[Y_t] = \lim_{n \to \infty} P[Y_{t_n}]
  \: .
\end{align*}
By uniform integrability (Lemma~\ref{lem:Submartingale.uniformIntegrable_of_antitone_of_ge}), we have that
\begin{align*}
  P[Y_{t+}] = \lim_{n \to \infty} P[Y_{t_n}]
  \: .
\end{align*}
Therefore $P[Y_{t+}] = P[Y_t]$, which concludes the proof.
\end{proof}


\begin{theorem}\label{thm:Martingale.exists_cadlag_modification}
  \uses{def:modification, def:Martingale, def:rightContinuous}
Let $X : T \to \Omega \to \mathbb{R}$ be a martingale with respect to a right-continuous filtration.
Then $X$ has a cadlag modification.
\end{theorem}

\begin{proof}
  \uses{thm:Submartingale.exists_cadlag_modification_iff_rightContinuous}
$X$ is in particular a submartingale, and for all $t \in T$, we have that $\mathbb{E}[X_t] = \mathbb{E}[X_0]$ by the martingale property.
Therefore $t \mapsto \mathbb{E}[X_t]$ is right-continuous, and we can apply Theorem~\ref{thm:Submartingale.exists_cadlag_modification_iff_rightContinuous}.
\end{proof}




\section{Cadlag modifications of (local) martingales}


TODO: this is partially or entirely redundant, consider removing this section.


% \begin{lemma}\label{lem:martingale_exists_dyadic_limit_left}
%   \uses{def:dyadics, def:Martingale}
%   Let $X=(X_t)_{t\in\mathcal{D}}$ be a martingale indexed by the dyadics. Then almost surely, for every $t\geq 0$ the limit
%   $$
%   \lim_{\stackrel{s\rightarrow t^-}{s\in\mathcal{D}}}X_s(\omega)
%   $$
%   exists and is finite.
% \end{lemma}

% \begin{proof}
%   See 8.2.1 of Pascucci.
% \end{proof}


% \begin{lemma}\label{lem:martingale_exists_dyadic_limit_right}
%   \uses{def:dyadics, def:Martingale}
%   Let $X=(X_t)_{t\in\mathcal{D}}$ be a martingale indexed by the dyadics. Then almost surely, for every $t\geq 0$ the limit
%   $$
%   \lim_{\stackrel{s\rightarrow t^+}{s\in\mathcal{D}}}X_s(\omega)
%   $$
%   exists and is finite.
% \end{lemma}

% \begin{proof}
%   See 8.2.1 of Pascucci.
% \end{proof}


% \begin{lemma}\label{lem:mg_is_cadlag}
%   \uses{def:hasUsualConditions, def:Martingale}
%   Let the filtered probability space satisfy the usual conditions.
%   Then every martingale $X$ admits a modification that is still a martingale with cadlag trajectories.
% \end{lemma}

% \begin{proof}
%   \uses{lem:martingale_exists_dyadic_limit_right,lem:martingale_exists_dyadic_limit_left}
%   See 8.2.3 of Pascucci.
% \end{proof}


\begin{lemma}\label{lem:exists_cadlag_mod_of_nonneg_submg}
  \uses{def:hasUsualConditions, def:Submartingale}
  Let the filtered probability space satisfy the usual conditions.
  Then every nonnegative submartingale $X$ admits a modification that is still a nonnegative submartingale with cadlag trajectories.
\end{lemma}

\begin{proof}
  %\uses{lem:martingale_exists_dyadic_limit_right,lem:martingale_exists_dyadic_limit_left}
  See 8.2.3 of Pascucci.
\end{proof}


\begin{lemma}\label{lem:exists_cadlag_mod_of_local_mg}
  \uses{def:hasUsualConditions}
  Let the filtered probability space satisfy the usual conditions.
  Then every local martingale $X$ admits a modification that is still a local martingale with cadlag trajectories.
\end{lemma}

\begin{proof}
  \uses{thm:Martingale.exists_cadlag_modification}
\end{proof}
