\chapter{Stochastic processes}
\label{chap:process}

Let $T$ be an index set and $\Omega$ a measurable space, with measure $\mathbb{P}$.
A stochastic process is a function $X : T \to \Omega \to E$, where $E$ is another measurable space, such that for all $t \in T$, $X_t : \Omega \to E$ is $\mathbb{P}$-a.e. measurable.


\begin{definition}[Law of a stochastic process]\label{def:processLaw}
  \leanok
The law of a stochastic process $X$ is the measure on the measurable space $E^T$ obtained by pushing forward the measure $\mathbb{P}$ by the map $\omega \mapsto X(\cdot, \omega)$.
\end{definition}

\textbf{Lean remark}: we don't use a Lean definition for the law, but write the map in full.

\begin{definition}[Modification]\label{def:modification}
  \leanok
We say that a stochastic process $Y$ is a \emph{modification} of another stochastic process $X$ if for all $t \in T$, $Y_t =_{\mathbb{P}\text{-a.e.}} X_t$.
\end{definition}

\textbf{Lean remark}: we don't use a Lean definition for being a modification, but write explicitly the condition $\forall t \in T,\ Y_t =_{\mathbb{P}\text{-a.e.}} X_t$~.

\begin{definition}[Indistinguishable]\label{def:indistinguishable}
  \leanok
We say that a stochastic processes $Y$ is a \emph{indistinguishable} from $X$ if $\mathbb{P}$-a.e., for all $t \in T$, $X_t = Y_t$.
\end{definition}

A summary of the next few lemmas is this:
\begin{itemize}
  \item indistinguishable $\implies$ modification $\implies$ same law,
  \item modification and continuous with $T$ separable $\implies$ indistinguishable.
\end{itemize}


\begin{lemma}\label{lem:Indistinguishable.Modification}
  \uses{def:indistinguishable, def:modification}
  \leanok
  \lean{modification_of_indistinduishable}
If $Y$ is indistinguishable from $X$, then $Y$ is a modification of $X$.
\end{lemma}

\begin{proof}\leanok
Obvious.
\end{proof}


\begin{lemma}\label{lem:map_eq_of_modification}
  \uses{def:modification}
  \leanok
  \lean{finite_distributions_eq}
Let $X, Y : T \to \Omega \to E$ be two stochastic processes that are modifications of each other.
Then for all $t_1, \ldots, t_n \in T$, the random vector $(X_{t_1}, \ldots, X_{t_n})$ has the same distribution as the random vector $(Y_{t_1}, \ldots, Y_{t_n})$.
That is, $X$ and $Y$ have same finite-dimensional distributions.
\end{lemma}

\begin{proof}\leanok
By the modification property, almost surely $X_{t_i} = Y_{t_i}$ for all $i \in [n]$.
Thus the function $\omega \mapsto (X_{t_1}(\omega), \ldots, X_{t_n}(\omega))$ is equal to $\omega \mapsto (Y_{t_1}(\omega), \ldots, Y_{t_n}(\omega))$ almost surely, hence the maps of $\mathbb{P}$ by these two functions are equal.
\end{proof}


\begin{lemma}\label{lem:map_eq_iff}
  \uses{def:processLaw}
  \leanok
  \lean{finite_distributions_eq_iff_same_law}
Let $X, Y : T \to \Omega \to E$ be two stochastic processes.
Then $X$ and $Y$ have same finite-dimensional distributions if and only if they have the same law.
\end{lemma}

\begin{proof}\leanok
TODO: consider the $\pi$-system of cylinder sets.
\end{proof}


\begin{lemma}\label{lem:indistinguishable_of_modification_of_continuous}
  \uses{def:modification, def:indistinguishable}
  \leanok
  \lean{indistinduishable_of_modification}
Let $T$ and $E$ be topological spaces and suppose that $T$ is separable Hausdorff.
Let $X, Y : T \to \Omega \to E$ be two stochastic processes that are modifications of each other and are almost surely continuous.
Then $X$ and $Y$ are indistinguishable.
\end{lemma}

\begin{proof}\leanok
Since $T$ is separable, it has a countable dense subset $D$.
Since $D$ is countable,
\begin{align*}
  (\forall t \in D, \mathbb{P}\text{-a.e.}, X_t = Y_t)
  \iff (\mathbb{P}\text{-a.e.}, \forall t \in D, X_t = Y_t)
\end{align*}
Hence by the modification property we have that almost surely, for all $t \in D$, $X_t = Y_t$.
Then almost surely $X$ and $Y$ are continuous functions which are equal on a dense subset of $T$: those two functions are equal everywhere.
\end{proof}
